---
title: "R Notebook"
output: html_notebook
---

```{r}
library(ggplot2)
library(forecast)
data <- read.csv("C:/Users/arush/Desktop/Spring 2020/Healthcare Analysis/HW/HW2/IE5400_HW02_Q7.csv")
data$Time <- c(1:78)
training <- data[1:68,]
validation <- data[69:78,]
ts <- ts(data = data, frequency = 1)
plot(ts)

```

```{r}
ggplot(data=data, mapping=aes(x=Time,y=Patient.Demand))+
geom_line()+
  geom_smooth(method = `lm`, se=FALSE)

```

From this graph, it can be clearly seen that there is an upward trend in the patient arrival, there is also a seasonality as a particular pattern can be seen being repeated over different time periods.

```{r}
#Dickey-Fuller test for stationarity
library(tseries)
adf.test(training$Patient.Demand, alternative="stationary",k=0)
#It can be seen that from the Dickey-Fuller test, since p<0.05, we can accept the hypothesis that the data is stationary

```

ACF, PACF
```{r}
acf(training$Patient.Demand, lag.max = 20)
pacf(training$Patient.Demand,lag.max = 20)
```

2)AR and MA Models
```{r}
#From PACF, let us take the lag at k=3 to build our AR model
arima(training$Patient.Demand, order = c(3,0,0))

#From ACF, let us take the lag at k=1 and build our MA Model
arima(training$Patient.Demand, order = c(0,0,1))

mymodel <- arima(training$Patient.Demand, order=c(3,0,1))
#We get the AIC as 589.48

#Fitting the model
res <- residuals(mymodel)
Box.test(res,lag = 20, type="Ljung-Box")

#It can be seen that the residuals have no correlation and the model is a good fit
```

3) Let us make use of the function auto.arima from the Forecast package in R to get the best (p,d,q) setting based on the least AIC and MLE
```{r}
ARIMA <- auto.arima(training$Patient.Demand, ic="aic",trace = T )

#We can see that the best setting is for (p,d,q)=(2,1,1) with AIC= 575.89. Thus, we will stick with this setting to forecast our values
```

Forecasting
```{r}
fc <- forecast(ARIMA,h=10)
predictions <- data.frame("Actual"=validation$Patient.Demand,"Predicted"=fc$mean,"Error"=validation$Patient.Demand-fc$mean)
predictions

accuracy(fc,validation$Patient.Demand)
```

CMORT Dataset
```{r}
data <- read.csv("cmort.csv", header = F)
names(data) <- "Patient_Arrival"
data$Time <- 1:500
training <- data[1:450,]
validation <- data[451:500,]
ts <- ts(data = data, frequency = 1)
plot(ts)

```

```{r}
ggplot(data=data, mapping=aes(x=Time,y=Patient_Arrival))+
geom_line()+
  geom_smooth(method = `lm`, se=FALSE)

#It can be observed from the plot that there exists a downaward trend of patient arrival and there also exists a seasonality which peaks at a certain time and then falls down and again rises up to another peak over a period of time
```

Dickey-Fuller test for stationarity
```{r}
 adf.test(training$Patient_Arrival, alternative="stationary",k=0)
#It can be seen that the p-value is less than 0.05, indicating that the data is not stationary
```

ACF,PACF
```{r}
acf(training$Patient_Arrival, lag.max = 20)
pacf(training$Patient_Arrival,lag.max = 20)
```

2)AR and MA Models
```{r}
#From PACF, let us take the lag at k=1 to build our AR model
arima(training$Patient_Arrival, order = c(1,0,0))

#From ACF, let us take the lag at k=2 and build our MA Model
arima(training$Patient_Arrival, order = c(0,0,2))

#We get an AIC of 2860.54

#Fitting the model
res <- residuals(mymodel)
Box.test(res,lag = 20, type="Ljung-Box")
#It can be seen that the residuals have no correlation and the model is a good fit
```

Let us make use of the function auto.arima from the Forecast package in R to get the best (p,d,q) setting based on the least AIC and MLE
```{r}
ARIMA <- auto.arima(training$Patient_Arrival, ic="aic",trace = T )
#We can see that the best setting is for (p,d,q)=(2,1,0) with AIC= 2859.09. Thus, we will stick with this setting to forecast our values
```

Forecasting
```{r}
fc <- forecast(ARIMA,h=50)
predictions <- data.frame("Actual"=validation$Patient_Arrival,"Predicted"=fc$mean,"Error"=validation$Patient_Arrival-fc$mean)
head(predictions)
accuracy(fc,validation$Patient_Arrival)
```

